Eliminating object prior-bias from sparse-projection tomographic reconstructions

### 2021-10-02 paper/post_tci.tex
Changes in text of RMSE

In this section, we first illustrate the advantage of using multiple
types of eigenspaces for the computation of weights map, and then
present reconstruction results on all datasets. For quantitative
evaluation, we compute the Structural Similarity Index Metric
(SSIM)~\cite{ssim}. Here, we assign higher importance to the
preservation of structure in a reconstruction by assigning the
following weights to structure, contrast and intensity respectively:
$0.7,0.2,0.1$. The reason for emphasizing the structure with an
exponent of 0.7 steps from the optimization process since the
intensity span and, in general, the histogram of reconstructed volumes
differ across various methods and solvers.

@@

\caption{SSIM of the reconstructions shown in Fig.~\ref{fig:reconstructions_diff_methods}:(a)-(g). The SSIM of ground-truth (ideal reconstruction) is 1.}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|}

@@

We observe
that our method reconstructs new structures while simultaneously
reducing streak artefacts. Table~\ref{table:potato_ssim} shows SSIM of
the reconstructed new regions using various methods.
\begin{comment}
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{blue}{blue}{\includegraphics[width=0.22\columnwidth]{../images/potato/post_tci/6_views/test_zoomed.png}}}
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/fbp_zoomed.png}}
\subcaptionbox{TV, no prior}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/tv_zoomed.png}}
\subcaptionbox{This paper}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/weighted_pca_zoomed.png}}
\caption{A zoomed-in version of the regions around the RoI of the reconstructions shown in Fig.~\ref{fig:potato_2D_results}.}
\label{fig:potato_zoomed_2D_results}
\end{figure}
\end{comment}
%-------------------------------------------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM of reconstructed potato from various
    methods. SSIM of ground-truth (ideal reconstruction) is 1.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline \textbf{2D} & 0.52
& 0.92 & \textcolor{red}{0.97} \\ \hline \textbf{3D} & 0.58 & 0.75 & \textcolor{red}{0.84}
\\ \hline
\end{tabular}
\label{table:potato_ssim}
\end{table}

%--------------------------Okra-- results
\subsubsection{\textbf{Okra}}
\label{Sec:okra_spatially_varying}
The test volume was reconstructed from a partial set of $6.7\%$ of projection views from which
ground truth was reconstructed (refer Sec.~\ref{sec:datasets}). 2D reconstructions of one of the
slices is shown in Fig.~\ref{fig:okra_2D_results}. A complete 3D
reconstruction can be found in in the supplementary material
~\cite{supp_paper}. The red RoI spans 7 consecutive slices where the
test is different from all of the previously scanned objects. %A
%zoomed-in region around the major region of change (red RoI) is shown
%in Fig.~\ref{fig:okra_zoomed_2D_results}.
As in the test, the
reconstruction by spatially-varying method shows the absence of the
deformity and better removal of sub-sampling artefacts when compared
to FDK and TV. This is also seen in the SSIM values in
Table~\ref{table:okra_ssim}.

@@

%---------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM of reconstructed okra from various
    methods. SSIM of ground-truth (ideal reconstruction) is 1.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline \textbf{2D} & 0.46
& 0.50 & \textcolor{red}{0.74} \\ \hline \textbf{3D} & 0.69 & 0.74 & \textcolor{red}{0.78}
\\ \hline
\end{tabular}
\label{table:okra_ssim}
\end{table}

@@

\begin{table}[!h]
  \centering
  \caption{SSIM of 2D reconstruction of radio-frequency ablation data from various methods. Values have been computed within the RoI. For ground-truth (ideal reconstruction), SSIM equals 1.}
\begin{tabular}{|l|c|c|c|c|}
\hline \textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline  0.73
& 0.91 & \textcolor{red}{0.95} \\ \hline 
\end{tabular}
\label{table:tmh_ssim}
\end{table}
%-----------------------------------------------------------------

@@

%----------------------------------------------------
\begin{comment}
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{yellow}{yellow}{\includegraphics[width=0.22\columnwidth]{../images/sprouts/post_TCI/2D/20_views/test_zoomed.png}}}
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/fbp_zoomed.png}}
\subcaptionbox{TV, no prior}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/tv_zoomed.png}}
%\subcaptionbox{\mbox{Uniform}\\prior}{\includegraphics[width=0.19\columnwidth]{../images/okra/pca_cropped.png}}\hfill
\subcaptionbox{This paper}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/weighted_pca_zoomed.png}}
\caption{A zoomed-in version of the regions around the RoI of the reconstructions shown in Fig.~\ref{fig:sprouts_2D_results}}
\label{fig:sprouts_zoomed_2D_results}
\end{figure}
\end{comment}
%--------------------------------------------------------------
\begin{table}[!h]
  \centering
  \caption{SSIM of reconstructed sprouts from various
    methods. The SSIM of ground-truth (ideal reconstruction) is 1.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} & \textbf{This paper}
 \\ \hline \textbf{2D} & 0.67
& 0.82 & \textcolor{red}{0.86} \\ \hline \textbf{3D} & 0.67 & 0.73 & \textcolor{red}{0.82}
\\ \hline
\end{tabular}
\label{table:sprouts_ssim}
\end{table}

@@

Most of the computation time in our technique is used for the
generation of low quality eigen-space by reconstructing each of the
templates using TV and FDK. If the set of projection views for every
test is fixed a priori, a reasonable assumption,
the set of low quality reconstructions can be
produced offline and stored in order to save computational costs. With
this assumption, Table~\ref{table:compute_time} shows the computation
times for various stages of reconstruction of a 2D slice of Liver
(refer code in~\cite{codeRepo}) on MATLAB 2019a on an AMD 2920X
12-Core Processor machine with 64GB RAM. If however, new sets of
projection views for every test are to be allowed, the low quality
reconstructions can still be performed efficiently using
parallelization since these reconstructions are completely
independent. Here below, we describe the time complexity of
spatially-varying prior technique.  Let
\begin{itemize}
\item $N$ denote the number of voxels of test volume,
\item $L$ denote the number of templates available,
\item $M$ denote the number of pilot reconstruction algorithms used for the computation of weights map, and
\item $F$ denote the time taken for optimization.
  $F$ consists of time taken for alternate minimization of Eq.~\ref{eq:spatially_varying_prior}, which in turn consists of time taken for repeated cycles of a closed-form update of $\boldsymbol{\alpha}$ (Eq.~\ref{eq:compute_alpha}) and an iterative optimization of  $\boldsymbol{x}$ (Eq.~\ref{eq:compute_x}). 
\end{itemize}
We ntoe that in practice $L$ is independent of $N$ and is likely to be $O(1)$  (e.g. about 12 scans in a year for disease prognosis).  Likewise $M$ is also $O(1)$. The speed of iterative optimization is dependent on the number of imaging views, and is profoundly impacted by the actual method used.  
Therefore, the time computational complexity of our method is sum of
the following
\begin{enumerate}
\item $O(NL^2+L^3)$ for creating eigenspace, which is, in practice linear, i.e., $O(N)$
\item $O(NL)*M$ for deriving weights map, which is, in practice linear, i.e., $O(N)$
\item and the time $F$  taken for iterative minimization, which as mentioned above is not the key contribution of this paper. 
\end{enumerate}
In short, the additional burden of the key idea of this paper viz. generating weights map is linear in the number of voxels, although admittedly the burden $F$ of repated optimization drives up the cost.

@@

\begin{table}[]
 \centering
  \caption{Computation times (in seconds) for various stages of reconstruction of a 2D slice of Liver (refer code in~\cite{codeRepo}) on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. Here we assume that low quality eigen-spaces of the templates are computed offline.}
\begin{tabular}{|l|l|l|l|}
\hline
Backprojection             & TV                        & \begin{tabular}[c]{@{}l@{}}Computation of \\ weights map\end{tabular} & \begin{tabular}[c]{@{}l@{}}Final reconstruction\\ with spatially-varying \\ prior\end{tabular} \\ \hline
\multicolumn{1}{|c|}{0.40} & \multicolumn{1}{c|}{8.44} & \multicolumn{1}{c|}{0.04}                                             & \multicolumn{1}{c|}{46.73}                                                                     \\ \hline
\end{tabular}
\label{table:compute_time}
\end{table}
\subsection{Requirements for applying deep learning techniques}
Deep learning techniques are powerful tools to learn an unknown
parameteric function relating an input space and an output space. In such
settings, our current goal of detecting new changes is a `prediction'
problem in a continuous solution space i.e., \textit{``given
  intensities of a voxel at various time instants in a longitudinal
  setting and partial measurements of the voxel at the current time,
  what will be the intensity of the same voxel at the current time
  instant?''}. This estimation can be learnt by a deep neural network
if there are hundreds of labeled data.  However, generalization of
this across multiple datasets poses a question mark.

>>>>

In this section, we first illustrate the advantage of using multiple
types of eigenspaces for the computation of weights map, and then
present reconstruction results on all datasets. For quantitative
evaluation, we compute the Structural Similarity Index Metric
(SSIM)~\cite{ssim}. Here, we assign higher importance to the
preservation of structure in a reconstruction by assigning the
following weights to structure, contrast and intensity respectively:
$0.7,0.2,0.1$. The reason for emphasizing the structure with an
exponent of 0.7 stems from the optimization process since the
intensity span (and, in general, the histogram) of reconstructed volumes
differ across various methods and solvers.

@@

\caption{SSIM and Relative MSE of the reconstructions shown in Fig.~\ref{fig:reconstructions_diff_methods}:(a)-(g). The SSIM of ideal reconstruction is 1.}
\begin{tabular}{|l|c|c|c|c|c|c|c|c|c|}

@@

We observe
that our method reconstructs new structures while simultaneously
reducing streak artefacts. Table~\ref{table:potato_ssim} shows SSIM of
the reconstructed new regions using various methods.
\begin{comment}
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{blue}{blue}{\includegraphics[width=0.22\columnwidth]{../images/potato/post_tci/6_views/test_zoomed.png}}}
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/fbp_zoomed.png}}
\subcaptionbox{TV, no prior}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/tv_zoomed.png}}
\subcaptionbox{This paper}{\includegraphics[width=0.23\columnwidth]{../images/potato/post_tci/6_views/weighted_pca_zoomed.png}}
\caption{A zoomed-in version of the regions around the RoI of the reconstructions shown in Fig.~\ref{fig:potato_2D_results}.}
\label{fig:potato_zoomed_2D_results}
\end{figure}
\end{comment}
%-------------------------------------------------------------------------
\begin{table}[!h]
  \centering
  \caption{The SSIM (Relative MSE) of reconstructed potato from various
    methods. The SSIM of ideal reconstruction is 1.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline \textbf{2D} & 0.52 / 0.16
& 0.92 / 0.00 & \textcolor{red}{0.97 / 0.00} \\ \hline \textbf{3D} & 0.58 / 0.29 & 0.75 / 0.24 & \textcolor{red}{0.84 / 0.18}
\\ \hline
\end{tabular}
\label{table:potato_ssim}
\end{table}

%--------------------------Okra-- results
\subsubsection{\textbf{Okra}}
\label{Sec:okra_spatially_varying}
The test volume was reconstructed from a partial set of $10\%$ of projection views from which
ground truth was reconstructed (refer Sec.~\ref{sec:datasets}). 2D reconstructions of one of the
slices is shown in Fig.~\ref{fig:okra_2D_results}. A complete 3D
reconstruction can be found in in the supplementary material
~\cite{supp_paper}. The red RoI spans 7 consecutive slices where the
test is different from all of the previously scanned objects. %A
%zoomed-in region around the major region of change (red RoI) is shown
%in Fig.~\ref{fig:okra_zoomed_2D_results}.
As in the test, the
reconstruction by spatially-varying method shows the absence of the
deformity and better removal of sub-sampling artefacts when compared
to FDK and TV. This is also seen in the SSIM values in
Table~\ref{table:okra_ssim}.

@@

%---------------------------------------
\begin{table}[!h]
  \centering
  \caption{The SSIM (Relative MSE) of reconstructed okra from various
    methods. SSIM of the ideal reconstruction is 1.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline \textbf{2D} & 0.46 / 0.22
& 0.50 / 0.16 & \textcolor{red}{0.74 / 0.06} \\ \hline \textbf{3D} & 0.65 / 0.35 & 0.70 / 0.32 & \textcolor{red}{0.71 / 0.29}
\\ \hline
\end{tabular}
\label{table:okra_ssim}
\end{table}

@@

\begin{table}[!h]
  \centering
  \caption{The SSIM (Relative MSE) of 2D reconstruction of radio-frequency ablation data from various methods. Values have been computed within the RoI. For ideal reconstruction, SSIM equals 1.}
\begin{tabular}{|l|c|c|c|c|}
\hline \textbf{Backprojection} & \textbf{TV} &
\textbf{This paper} \\ \hline  0.73 / 0.05
& 0.91 / 0.17 & \textcolor{red}{0.95 / 0.03} \\ \hline 
\end{tabular}
\label{table:tmh_ssim}
\end{table}
%-----------------------------------------------------------------

@@

%----------------------------------------------------
\begin{comment}
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{yellow}{yellow}{\includegraphics[width=0.22\columnwidth]{../images/sprouts/post_TCI/2D/20_views/test_zoomed.png}}}
\subcaptionbox{FDK, no prior}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/fbp_zoomed.png}}
\subcaptionbox{TV, no prior}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/tv_zoomed.png}}
%\subcaptionbox{\mbox{Uniform}\\prior}{\includegraphics[width=0.19\columnwidth]{../images/okra/pca_cropped.png}}\hfill
\subcaptionbox{This paper}{\includegraphics[width=0.23\columnwidth]{../images/sprouts/post_TCI/2D/20_views/weighted_pca_zoomed.png}}
\caption{A zoomed-in version of the regions around the RoI of the reconstructions shown in Fig.~\ref{fig:sprouts_2D_results}}
\label{fig:sprouts_zoomed_2D_results}
\end{figure}
\end{comment}
%--------------------------------------------------------------
\begin{table}[!h]
  \centering
  \caption{The SSIM (Relative MSE) of reconstructed sprouts from various
    methods. The SSIM of ideal reconstruction) is 1.}
\begin{tabular}{|l|c|c|c|c|c|}
\hline &
\textbf{Backprojection} & \textbf{TV} & \textbf{This paper}
 \\ \hline \textbf{2D} & 0.67 / 0.23
& 0.82 / 0.12 & \textcolor{red}{0.86 / 0.07} \\ \hline \textbf{3D} & 0.67 / 0.53 & 0.73 / 0.48 & \textcolor{red}{0.82 / 0.30}
\\ \hline
\end{tabular}
\label{table:sprouts_ssim}
\end{table}

@@

Most of the computation time in our technique is used for the
generation of low quality eigen-space by reconstructing each of the
templates using TV and FDK. If the set of projection views for every
test is fixed a priori, a reasonable assumption,
the set of low quality reconstructions can be
produced offline and stored in order to save computational costs. With
this assumption, Table~\ref{table:compute_time} shows the computation
times for various stages of reconstruction of a 2D slice of Liver
(refer code in~\cite{codeRepo}) on MATLAB 2019a on an AMD 2920X
12-Core Processor machine with 64GB RAM. If however, new sets of
projection views for every test are to be allowed, the low quality
reconstructions can still be performed efficiently using
parallelization since these reconstructions are completely
independent. Here below, we describe the time complexity of
spatially-varying prior technique.  Let
\begin{itemize}
\item $N$ denote the number of voxels of test volume,
\item $L$ denote the number of templates available,
\item $M$ denote the number of pilot reconstruction algorithms used for the computation of weights map, and
\item $F$ denote the time taken for optimization.
  $F$ consists of time taken for alternate minimization of Eq.~\ref{eq:spatially_varying_prior}, which in turn consists of time taken for repeated cycles of a closed-form update of $\boldsymbol{\alpha}$ (Eq.~\ref{eq:compute_alpha}) and an iterative optimization of  $\boldsymbol{x}$ (Eq.~\ref{eq:compute_x}). 
\end{itemize}
We note that in practice $L$ is independent of $N$ and is likely to be
$O(1)$  (e.g. about 12 scans in a year for disease prognosis).
Likewise $M$ is also $O(1)$. The speed of the iterative optimization is dependent on the number of imaging views, and is profoundly impacted by the actual numerical optimization technique used.  
Therefore, the time computational complexity of our method is sum of
the following
\begin{enumerate}
\item $O(NL^2+L^3)$ for creating eigenspace, which is, in practice linear, i.e., $O(N)$
\item $O(NLM)$ for deriving weights map, which is, in practice linear, i.e., $O(N)$
\item and the time $F$  taken for iterative minimization, which as mentioned above is not the key contribution of this paper. 
\end{enumerate}
In short, the additional burden of the key idea of this paper
viz. generating weights map is linear in the number of voxels,
although admittedly the burden $F$ of the alternate minimization procedure drives up
the cost.

@@

\begin{table}[]
 \centering
  \caption{Computation times (in seconds) for various stages of reconstruction of a 2D slice of Liver (refer code in~\cite{codeRepo}) on MATLAB 2019a on an  AMD 2920X 12-Core Processor machine with 64GB RAM. Here we assume that low quality eigen-spaces of the templates are computed offline.}
\begin{tabular}{|l|l|l|l|}
\hline
 \begin{tabular}[c]{@{}l@{}} FBP (Sufficient \\ views)\end{tabular}             & \begin{tabular}[c]{@{}l@{}} TV (Moderate \\ views)  \end{tabular}                      & \begin{tabular}[c]{@{}l@{}}Computation of \\ weights map \\ (Very few \\views)\end{tabular} & \begin{tabular}[c]{@{}l@{}}Alternate\\ Minimization \\ (Very few \\views)\end{tabular} \\ \hline
\multicolumn{1}{|c|}{0.01} & \multicolumn{1}{c|}{6.00} & \multicolumn{1}{c|}{0.01}                                             & \multicolumn{1}{c|}{34.00}                                                                     \\ \hline
\end{tabular}
\label{table:compute_time}
\end{table}
\subsection{Requirements for applying deep learning techniques}
Deep learning techniques are powerful tools to learn an unknown
parameteric function relating an input space and an output space. In such
settings, our current goal of detecting new changes is a `prediction'
problem in a continuous solution space i.e., \textit{``given
  intensities of a voxel at various time instants in a longitudinal
  setting and partial measurements of the voxel at the current time,
  what will be the intensity of the same voxel at the current time
  instant?''}. This estimation can be learnt by a deep neural network
if there are hundreds of labeled data.  However, generalization of
this across multiple datasets poses a question mark.
